Recognizing Strong Gravitational Lenses

### 2015-03-16 doc/Course_Paper/paper.tex
updating code. added one more reference. pushing final paper tex and pdf

% Discuss how well things went. Or didn't go.
The need for new automated detection algorithms for finding strong
gravitational lenses will only become more pressing in the next decade, as it
becomes infeasible for scientists to scan images by eye for such systems. Using
the \textsc{Space Warps} dataset, we have examined how convolution neural
networks trained both on this particular dataset and on other datasets can
perform at the detection task.
% TODO: brief blurb about our CNN
We find that features extracted from a convolution neural network trained on
the classification of galaxy morphology (with a linear support vector
classifier on top for converting the feature vector to a binary ``lens'' and
``not-lens'') performs admirably but markedly worse than the citizen-scientists
trained on the dataset. Further work examining improvements by retraining the
whole neural network could lead to a generic classification machine that takes
images from \textit{any} galaxy survey and states whether the image contains a
strong lens or not. Additionally there is much potential in both direct
convolution neural networks and transfer learning from other networks in
linking the classification outputs of the networks with the probability
estimates of the \textsc{Space Warps} system, which also examined several
thousands more ``unknown'' systems and could lead to more gravitational lenses
being identified.

@@

% TODO: copy relevant bib stuff over so I don't have a 4 mb bib
{\small
\bibliographystyle{ieee}
\bibliography{coursebib}
}

>>>>

% Discuss how well things went. Or didn't go.
The need for new automated detection algorithms for finding strong
gravitational lenses will only become more pressing in the next decade, as it
becomes infeasible for scientists to scan images by eye for such systems. Using
the \textsc{Space Warps} dataset, we have examined how convolution neural
networks trained both on this particular dataset and on other datasets can
perform at the detection task.

While our Convolutional Neural Net trained on just the \textsc{Space Warps} dataset 
has not yet outperformed citizen-scientists at sufficiently low false positive rates, there is 
ample reason to believe marked improvements can be made with more time and 
CPU power. In addition to training on more images (and augmented images) for longer, 
we can expect to gain a significant reduction in training time by taking advantage of 
\textsc{theano}'s GPU capabilities and by implementing batch normalization. \cite{Ioffe:2015aa} 
It may also be beneficial to explore larger and more expressive CNN architectures.

In our application of Transfer Learning, we find that features extracted from a convolution neural network trained on
the classification of galaxy morphology (with a linear support vector
classifier on top for converting the feature vector to a binary ``lens'' and
``not-lens'') performs admirably but markedly worse than the citizen-scientists
trained on the dataset. Further work examining improvements by retraining the
whole neural network could lead to a generic classification machine that takes
images from \textit{any} galaxy survey and states whether the image contains a
strong lens or not. Additionally there is much potential in both direct
convolution neural networks and transfer learning from other networks in
linking the classification outputs of the networks with the probability
estimates of the \textsc{Space Warps} system, which also examined several
thousands more ``unknown'' systems and could lead to more gravitational lenses
being identified.

@@

{\small
\bibliographystyle{ieee}
\bibliography{coursebib}
}
